Sperm should evolve to make female meiosis fair.

### 2014-07-25 writeup_and_notes/GeneticsResub/reviews.tex
changes to reviews

\begin{document}
Please find attached a new manuscript for your evaluation for publication in GENETICS, titled  `Sperm should evolve to make female meiosis fair.' 
The positive and constructive reviews of our previous submission with the same title (\emph{MS ID\#: GENETICS/2014/167064})  where extremely useful in our crafting of this revision. 
Below we discuss how this paper was improved by addressing, in detail, previous concerns \emph{point-by-point}, but first we address the major  concerns with our previous submission.
\begin{itemize}
\begin{singlespace}
\item \emph{{\bf{Major concern 1 -- }} The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis. However, as Reviewer 1 points out, this is different from showing that factors expressed in females should evolve to delay completion of meiosis until after fertilization. There is general agreement that the argument needs to be completed so that it supports all conclusions and that the justification for study of the question be improved. Reviewer 2 has provided many valuable suggestions along these lines.}  \newline \\
{\bf{Response -- }} This revision makes it clear that our primary argument is that, given a chance to influence female meiosis, sperm will evolve to increase its fairness (as reflected in our title), {\bf{AND NOT}} that females should delay the completion of meiosis so as to recruit help from sperm in making meiosis fair (as could have been interpreted from our previous draft). 
\end{singlespace}

\begin{singlespace}
\item \emph{{\bf{Major concern 2 -- }} The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.). As a result, it is unclear whether GENETICS is the appropriate or even the intended forum for this work.} \newline \\
{\bf{Response -- }}  We believe our manuscript presents 
(1) a rigorous and theoretical model, and 
(2) research questions that can be addressed by molecular genetics, transmission genetics, studies of genome structure, phylogenetics and population genetics. As such, GENETICS is a particularly appropriate venue for our paper. 
Our view is echoed by comments from Reviewer 2 who stated  `Overall I found the paper to be well written and organized, and because of this, the theory that the authors describe will be accessible to a broad spectrum of the readership of Genetics ? not just other theoreticians.'
The suitability of the topic of our manuscript is clear from our reference list -- a strong plurality (over 25\%) of our works cited appeared in GENETICS, placing our work squarely within the tradition of GENETICS papers.
We have therefor corrected issues with our formatting, and included a METHODS section to introduce our models, as is the tradition in GENETICS. \\
\end{singlespace} 
\end{itemize}
We thank you for your previous comments and for consideration of this revised manuscript. 
To aid in your consideration, we provide  previous reviews and a description of how we have addressed individual comments, below. 
\\
\\
Sincerely,  \\
Yaniv Brandvain and Graham Coop.

@@

{\bf{Response:}} We appreciate that the Reviewer found the paper to be `a pleasure 
to read' and that the models  where `straightforward and clearly explained.'
We do however insist that this is a sufficient advance for publication in GENETICS --
 we illustrate the expected outcomes of a previously under-appreciated feature of sperm-
 egg interactions and present a series of novel and testable hypotheses. 
\begin{enumerate}
\item
\begin{singlespace}
\emph{A major part of the motivation for the model seems to be to explain why, in
animals, meiosis is typically suspended until fertilization occurs (providing
the opportunity for males to influence female drive).  This seems a poor
motivation for the project because it would require a bizarre form of logic by
which the evolution of this delay in meiosis evolved with the ``foresight'' that
there would then be subsequent evolution of male influence on female drive. 
More directly, if the authors' interest is in the evolution of arrest in female
meiosis, then it would be better to do a model that explicitly addressed that
issue.  }

@@

\begin{singlespace}
{\bf{Response -- }} We thank the reviewer for this comment, and have reworked our manuscript to focus 
	around our major result -- that sperm are expected to evolve to prevent female meiotic drive. 
	We completely agree that this delay connote be favored by ``evolutionary foresight'' and have 
	modified our manuscript to ensure that readers do not interpret our argument in this way 
	(in fact, we specifically clarify that this is not our argument in the revised introduction and discussion). 
	We view this as an important point, and as the significant change in this new manuscript. 
	As suggested by R1, we focus our discussion of the fertilization requirement to pointing out that it 
	provides sperm an opportunity to influence female meiosis.

	We clarify this point by referring readers to a list of hypotheses proposed to explain female meiotic arrest. 
	We point pout that, whatever favors this arrest,	 the threat of sperm collaboration in meiotic drive does not 
	represent a selective force favoring the  completion of meiosis before fertilization.
	We do argue, however, that whatever the evolutionary reason for the origin of this delay, the cost o losing 
	sperm protection against meiotic drive could play a role in the evolutionary maintenance  of the female meiotic arrest. \\
\end{singlespace}

@@

\begin{singlespace}
	{\bf{Response -- }}  We believe that this is an important phenomenon. 
	We have modified the manuscript to clarify that it is  quite impressive that we find evidence
	for male influence on female drive in BOTH cases where this phenomenon could possibly be observed.
	We note that although direct evidence for female meiotic drive is modest, 
	most experiments to dates have lacked the power to detect most modest drivers. 
	We believe that given our theoretical predictions and the affordability of large scale GBS experiments,
	data weighing in on this paper are on its way. 
	We appreciate the opportunity to get ahead of this data by offering strong predictions that could guide analyses.
\end{singlespace}
\end{enumerate}

@@

\begin{enumerate}
\begin{singlespace}
\item	\emph{The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
female meiotic drive.  In species like maize, where every chromosome has two or
more established meiotic drivers (``chromosomal knobs''), the importance of
meiotic drive is obvious, but in most species there is little or no evidence for
the current operation of meiotic drive -- although low levels of transmission
ratio distortion (TRD) are becoming increasingly well established from
genome-wide SNP studies.  However, the rapid divergence of telomeric sequences
among closely related species is certainly consistent with the widespread
operation of transient episodes of meiotic drive via centromere drive.  In the
present manuscript, the prevalence and importance of female meiotic drive is
quickly summarized by the terse statements ``Meiotic drivers observed in both
plants [5, 10, 9], and animals [1, 29, 21] highlight this conflict -- the selfish
benefits and the associated pleiotropic fitness costs of drive sustain a
balanced polymorphism [24], and often generate ongoing evolutionary escalations
of drive suppressors and enhancers [8, 9].''  I think the paper would be far more
convincing and appreciated if the authors put a tad more effort into convincing
the reader that female meiotic drive is sufficiently common to be a widespread
selective force.}
\end{singlespace}
\begin{singlespace}
	{\bf{Response -- }}  We agree and have added more support of the likely importance and prevalence of female meiotic drive. 
	We hope that new theoretical work (including ours) coupled with the falling expense of GBS, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
\end{singlespace}

>>>>

\begin{document}
Please find attached a new manuscript for your evaluation for publication in GENETICS, titled  `Sperm should evolve to make female meiosis fair.' 
The positive and constructive reviews of our previous submission with the same title (\emph{MS ID\#: GENETICS/2014/167064})  where extremely useful in our crafting of this revision. 
Below we discuss how this paper was improved by addressing, in detail, previous concerns \emph{point-by-point}, but first we address the major  concerns with our previous submission.
\begin{itemize}
\begin{singlespace}
\item \emph{{\bf{Major concern 1 -- }} The paper as written establishes that factors expressed in sperm should generally evolve to suppress drivers of meiotic drive in female meiosis. However, as Reviewer 1 points out, this is different from showing that factors expressed in females should evolve to delay completion of meiosis until after fertilization. There is general agreement that the argument needs to be completed so that it supports all conclusions and that the justification for study of the question be improved. Reviewer 2 has provided many valuable suggestions along these lines.}  \newline \\
{\bf{Response -- }} \gc{We thank the reviewers and editor for highlighting
this point. In this revision we make it clear} that our primary
argument is that, given a chance to influence female meiosis, sperm
will evolve to increase its fairness (as reflected in our title), {\it
  and not} that females should delay the completion of meiosis so as to recruit help from sperm in making meiosis fair (as could have been interpreted from our previous draft). 
\end{singlespace}

\begin{singlespace}
\item \emph{{\bf{Major concern 2 -- }} The format of this submission departs from the style of GENETICS (model description and analysis, literature citations, etc.). As a result, it is unclear whether GENETICS is the appropriate or even the intended forum for this work.} \newline \\
{\bf{Response -- }}  We believe our manuscript presents 
(1) a rigorous and theoretical model, and 
(2) research questions that can be addressed by molecular genetics,
transmission genetics, studies of genome structure, phylogenetics and
population genetics. As such, GENETICS is a particularly appropriate
venue for our paper.  Our view is echoed by comments from Reviewer 2 who stated  `Overall I found the paper to be well written and organized, and because of this, the theory that the authors describe will be accessible to a broad spectrum of the readership of Genetics not just other theoreticians.'
The suitability of the topic of our manuscript is clear from our reference list -- a strong plurality (over 25\%) of our works cited appeared in GENETICS, placing our work squarely within the tradition of GENETICS papers.
\gc{We apologize for the formatting issues. We have corrected these
  issues, and included a short METHODS section to introduce our models.} \\
\end{singlespace} 
\end{itemize}
We thank you for your previous comments and for consideration of this revised manuscript. 
To aid in your consideration, we provide \gc{the previous reviews and a
detailed response below. }
\\
\\
Sincerely,  \\
Yaniv Brandvain and Graham Coop.

@@

%%I'd not use quotes right below the quoted text, seems trite.
{\bf{Response:}} \gc{We appreciate that the Reviewer enjoyed the paper.}
We do however insist that this is a sufficient advance for publication in GENETICS --
 we illustrate the expected outcomes of a previously under-appreciated feature of sperm-
 egg interactions and present a series of novel and testable hypotheses. 
\begin{enumerate}
\item
\begin{singlespace}
\emph{A major part of the motivation for the model seems to be to explain why, in
animals, meiosis is typically suspended until fertilization occurs (providing
the opportunity for males to influence female drive).  This seems a poor
motivation for the project because it would require a bizarre form of logic by
which the evolution of this delay in meiosis evolved with the ``foresight'' that
there would then be subsequent evolution of male influence on female drive. 
More directly, if the authors' interest is in the evolution of arrest in female
meiosis, then it would be better to do a model that explicitly addressed that
issue.  }

@@

\begin{singlespace}
{\bf{Response -- }} We thank the reviewer for this comment, and have reworked our manuscript to focus 
	around our major result -- that sperm are expected to evolve to prevent female meiotic drive. 
	We completely agree that this delay \gc{cannot} be favored by ``evolutionary foresight'' and have 
	modified our manuscript to ensure that readers do not
        interpret our argument in this way. \gc{We have now added
          significantly to our  revised introduction and discussion to
          clarify this point. 	
We also now more explicitly refer readers to a discussion of
        hypotheses proposed to explain female meiotic arrest (Mira
        '98), so that our discussion can be seen as being distinct
        from that. }

%	As suggested by R1 \gc{???}, we focus our discussion of the fertilization requirement to pointing out that it 
%	provides sperm an opportunity to influence female meiosis.  
%	
%	We point out that, whatever favors this arrest, the threat of sperm collaboration in meiotic drive does not 
%	represent a selective force favoring the completion of meiosis before fertilization.
%	We do argue, however, that whatever the evolutionary reason for the origin of this delay, the cost o losing 
%	sperm protection against meiotic drive could play a role in the evolutionary maintenance of the female meiotic arrest. \\
\end{singlespace}

@@

\begin{singlespace}
	{\bf{Response -- }}  \gc{We believe that this is an important
        phenomenon. We have strengthened our discussion of this point
        throughout the introduction,
        we have also highlighted the points raised by R2 about 
        the evidence of pervassive female drive and the opportunities for
        sperm-based products
        to interfer with female meiosis.
	We have modified the manuscript to clarify that it is quite impressive that there is evidence
	for male influence on female drive at both of the two described
        female meiotic drive elements in animals. 
	We note that although direct evidence for polymorphic female meiotic drive
        in animals is modest (i.e. these two loci in mice), 
	this likely reflects the fact that it is very hard to study
        outside of model organisms. 
Given the increasing tractibility of genetics in many traditionally
        non-model systems, and the growing awareness of the importance
        of genetic conflicts in shaping molecular systems, we think
        that it is timely to offer a framework to understand one
        aspect of sperm-egg 
        interactions and provide
strong predictions that could guide analyses.}
\end{singlespace}
\end{enumerate}

@@

\begin{enumerate}
\begin{singlespace}
\item	\emph{The importance of the wok hinges critically on the the degree to which:  {\bf{\emph{i)}}}
female meiotic drive operates in nature, and {\bf{\emph{ii)}}} sperm influence the level of
female meiotic drive.  In species like maize, where every chromosome has two or
more established meiotic drivers (``chromosomal knobs''), the importance of
meiotic drive is obvious, but in most species there is little or no evidence for
the current operation of meiotic drive -- although low levels of transmission
ratio distortion (TRD) are becoming increasingly well established from
genome-wide SNP studies.  However, the rapid divergence of telomeric sequences
among closely related species is certainly consistent with the widespread
operation of transient episodes of meiotic drive via centromere drive.  In the
present manuscript, the prevalence and importance of female meiotic drive is
quickly summarized by the terse statements ``Meiotic drivers observed in both
plants [5, 10, 9], and animals [1, 29, 21] highlight this conflict -- the selfish
benefits and the associated pleiotropic fitness costs of drive sustain a
balanced polymorphism [24], and often generate ongoing evolutionary escalations
of drive suppressors and enhancers [8, 9].''  I think the paper would be far more
convincing and appreciated if the authors put a tad more effort into convincing
the reader that female meiotic drive is sufficiently common to be a widespread
selective force.}
\end{singlespace}
\begin{singlespace}
	{\bf{Response -- }}  We agree and have added more support of the likely importance and prevalence of female meiotic drive. 
	We hope that new theoretical work (including ours) coupled
        with the increasing easy of genetics in many system, will encourage the discovery of additional drive systems in non-model organisms in the coming years.
\end{singlespace}
